A survey of registration practices among observational researchers using preexisting datasets

### 2023-03-18 manuscript/interactive_report.Rmd
cosmetic changes to plot, small bug fixes with the code, styling

```{r, echo=FALSE}
inputPanel(
  sliderInput("n_studies", label = "Number of studies using ALSPAC:",
              min = min(processed$n_studies, na.rm = T),
              max = max(processed$n_studies, na.rm = T), value = c(0, max(processed$n_studies, na.rm = T)), step = 1),
  selectInput("methods", label = "Used the following methods:",
              choices = c(
                "Blind" = "method_blind",
                "Script" = "method_script",
                "Preregistration" = "method_preregistered",
                "Confirmatory" = "method_confirmatory", 
                "Exploratory" = "method_exploratory"),
              selected = NULL, multiple = TRUE),
  selectInput("concerned", label = "Level of concern with rigour and reproducibility:",
              choices = c(
                "Less concerned",
                "Neutral",
                "More concerned"),
              selected = c(
                "Less concerned",
                "Neutral",
                "More concerned"), multiple = TRUE),
  selectInput("language", label = "Programming language used for the analysis:",
              choices = c(
                "R",
                "Stata",
                "SPSS",
                "SAS",
                "Python",
                "Mplus",
                "Bash",
                "MATLAB",
                "Nextflow",
                "plink2",
                "irrelevant"
              ),
              selected = c(
                "R",
                "Stata",
                "SPSS",
                "SAS",
                "Python",
                "Mplus",
                "Bash",
                "MATLAB",
                "Nextflow",
                "plink2",
                "irrelevant"
              ), multiple = TRUE)
  )

# Filter data
filtered_data <- reactive({
  concerned_options <-list(
    "Less concerned" =  c("very much less concerned", "less concerned", "somewhat less concerned"),
    "Neutral" =  c("as concerned as a typical researcher in my field"),
    "More concerned" = c("somewhat more concerned", "more concerned", "Very much more concerned")
  )

@@

  processed %>% 
    filter(n_studies >= input$n_studies[1] & n_studies <= input$n_studies[2] | is.na(n_studies)) %>% 
    filter(if_all(input$methods, ~ . != "Never or almost never")) %>% 
    filter(concerned %in% concerned_filter | is.na(concerned)) %>% 
    filter(str_detect(programming_language, paste(input$language, collapse="|")) | is.na(programming_language))
  })
# select(filtered_data(), concerned)
# shiny::renderTable(filtered_data())
shiny::renderText({nrow(filtered_data())})
```

@@

```{r echo=FALSE}
# Concerned figure data preparation
concerned_plot_data <- reactive({
  processed %>% 
  select(concerned) %>% 
  mutate(
    concerned = tolower(concerned),
    concerned = factor(concerned, levels = c(
      "very much less concerned",
      "less concerned",
      "somewhat less concerned",
      "as concerned as a typical researcher in my field",
      "somewhat more concerned",
      "more concerned",
      "very much more concerned"
      )
    )
  ) %>% 
    filter(!is.na(concerned))
})

>>>>

```{r, echo=FALSE}
inputPanel(
  sliderInput("n_studies", label = "Number of studies using ALSPAC:",
              min = min(processed$n_studies, na.rm = T),
              max = max(processed$n_studies, na.rm = T), value = c(0, max(processed$n_studies, na.rm = T)), step = 1),
  selectInput("methods", label = "Used the following methods:",
              choices = list(
                "Blind" = list("Never or almost never", "Sometimes", "About half the time", "Most of the time", "Always or almost always", "Missing", "I don't understand the question"),
                "Script" = list("Never or almost never", "Sometimes", "About half the time", "Most of the time", "Always or almost always", "Missing", "I don't understand the question"),
                "Preregistration" = list("Never or almost never", "Sometimes", "About half the time", "Most of the time", "Always or almost always", "Missing", "I don't understand the question"),
                "Confirmatory" = list("Never or almost never", "Sometimes", "About half the time", "Most of the time", "Always or almost always", "Missing", "I don't understand the question"), 
                "Exploratory" = list("Never or almost never", "Sometimes", "About half the time", "Most of the time", "Always or almost always", "Missing", "I don't understand the question")),
              selected = NULL, multiple = TRUE),
  selectInput("concerned", label = "Level of concern with rigour and reproducibility:",
              choices = c(
                "Less concerned",
                "Neutral",
                "More concerned"),
              selected = c(
                "Less concerned",
                "Neutral",
                "More concerned"), multiple = TRUE),
  selectInput("language", label = "Programming language used for the analysis:",
              choices = c(
                "R",
                "Stata",
                "SPSS",
                "SAS",
                "Python",
                "Mplus",
                "Bash",
                "MATLAB",
                "Nextflow",
                "plink2",
                "irrelevant"
              ),
              selected = c(
                "R",
                "Stata",
                "SPSS",
                "SAS",
                "Python",
                "Mplus",
                "Bash",
                "MATLAB",
                "Nextflow",
                "plink2",
                "irrelevant"
              ), multiple = TRUE)
  )

# Filter data
filtered_data <- reactive({
  # list(
  #               "Blind" = "method_blind",
  #               "Script" = "method_script",
  #               "Preregistration" = "method_preregistered",
  #               "Confirmatory" = "method_confirmatory", 
  #               "Exploratory" = "method_exploratory")
  # 
  concerned_options <-list(
    "Less concerned" =  c("very much less concerned", "less concerned", "somewhat less concerned"),
    "Neutral" =  c("as concerned as a typical researcher in my field"),
    "More concerned" = c("somewhat more concerned", "more concerned", "Very much more concerned")
  )

@@

  processed %>% 
    filter(n_studies >= input$n_studies[1] & n_studies <= input$n_studies[2] | is.na(n_studies)) %>% 
    # filter(if_all(input$methods, ~ . != "Never or almost never")) %>% 
    filter(concerned %in% concerned_filter | is.na(concerned)) %>% 
    filter(str_detect(programming_language, paste(input$language, collapse="|")) | is.na(programming_language))
  })
# select(filtered_data(), concerned)
# shiny::renderTable(filtered_data())
shiny::renderText({input$methods})
```

@@

```{r echo=FALSE}
# Concerned figure data preparation
concerned_plot_data <- reactive({
  filtered_data() %>% 
  select(concerned) %>% 
  mutate(
    concerned = tolower(concerned),
    concerned = factor(concerned, levels = c(
      "very much less concerned",
      "less concerned",
      "somewhat less concerned",
      "as concerned as a typical researcher in my field",
      "somewhat more concerned",
      "more concerned",
      "very much more concerned"
      )
    )
  ) %>% 
    filter(!is.na(concerned))
})

### 2023-03-18 manuscript/manuscript.Rmd
cosmetic changes to plot, small bug fixes with the code, styling

```{r setup, include = FALSE, warning=FALSE, message=FALSE}
library("papaja")
library(tidyverse)
library(here)
library(kableExtra)
library(gt)
library(patchwork)
# The developer version is needed
library(likert)
library(knitr)
library(viridis)
# r_refs("references.bib")
```

```{r custom_scripts}
# Source R scripts
r_scripts <- list.files(here::here("R/"), full.names = TRUE)
purrr::walk(r_scripts, source)
```

```{r preprocessing, include=FALSE}
# loads source data, performs preprocessing, saves raw datafiles
xfun::Rscript_call(
  rmarkdown::render,
  list(
    input = here::here("preprocessing", "ecaw_source_raw_preprocessing.Rmd"),
    output_format = 'html_document')
  )

# loads raw data, performs preprocessing, saves processed datafiles
xfun::Rscript_call(
  rmarkdown::render,
  list(
    input = here::here("preprocessing", "ecaw_raw_processed_preprocessing.Rmd"),
    output_format = 'html_document')
  )
```

```{r loading_data}
processed <- read_csv(here::here("data/processed/ecaw_processed_data.csv"))
```

```{r analysis_preferences}
# Seed for random number generation
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

@@

```{r include=FALSE}
# List of used programming languages and their counts text
language_text <-
  processed %>% 
  rename(lang = programming_language) %>% 
  separate_rows(lang, sep = ",") %>% 
  count(lang) %>% 
  filter(!is.na(lang)) %>% 
  arrange(desc(n)) %>% 
  # Removing irrelevant
  filter(lang != "irrelevant") %>% 
  mutate(lang_n = glue::glue("{lang} (n = {n})")) %>% 
  pull(lang_n) %>% 
  glue::glue_collapse(., sep = ", ", last = ", and ")

# Calculate concerned percentage
concerned <- support_percentage(
  processed,
  concerned,
  c(
    "very much less concerned",
    "less concerned",
    "somewhat less concerned",
    "as concerned as a typical researcher in my field",
    "somewhat more concerned",
    "more concerned",
    # This level label starts with a capital letter
    "Very much more concerned"
    )
  )
```

@@

```{r include=FALSE}
# Calculate typical trustworthiness and reproducibility ratings
typically_trustworthy <- support_percentage(
  processed,
  typically_trustworthy,
  c(
    "Strongly disagree",
    "Somewhat disagree",
    "Neither agree nor disagree",
    "Somewhat agree",
    "Strongly agree"),
  "I don't understand the question"
  )

typically_reproducible <- support_percentage(
  processed,
  typically_reproducible,
  c(
    "Strongly disagree",
    "Somewhat disagree",
    "Neither agree nor disagree",
    "Somewhat agree",
    "Strongly agree"),
  "I don't understand the question"
  )

# Calculate ecaw trustworthiness and reproducibility ratings
ecaw_trustworthy <- support_percentage(
  processed,
  ecaw_trustworthy,
  c(
    "Much less",
    "Somewhat less",
    "About the same",
    "Somewhat more",
    "Much more"),
  "I don't understand the question"
  )

ecaw_reproducible <- support_percentage(
  processed,
  ecaw_reproducible,
  c(
    "Much less",
    "Somewhat less",
    "About the same",
    "Somewhat more",
    "Much more"),
  "I don't understand the question"
  )
```

```{r typicallyEcawPlot, warning=FALSE, fig.cap="Caption goes here...", fig.align="center", out.width="100%", fig.width=12.2, fig.height=6}
# Create typically trustworthy and reproducible plot
# Check if there is any missing and don't understand responses separately
# processed %>%
#   select(contains("typically_")) %>%
#   dplyr::filter(is.na(typically_reproducible) | is.na(typically_trustworthy) | typically_reproducible == "I don't understand the question" | typically_trustworthy == "I don't understand the question") %>%
#   nrow()

# Prepare plot data
typically_plot_data <- 
  processed %>%
  select(contains("typically_")) %>% 
  mutate(
    across(
      .fns = ~ factor(., levels =   c(
    "Strongly disagree",
    "Somewhat disagree",
    "Neither agree nor disagree",
    "Somewhat agree",
    "Strongly agree"))
    )) %>% 
  rename(
    Reproducible = typically_reproducible,
    Trustworthy = typically_trustworthy
  )

@@

# Create figure
typically_plot <- 
  plot(typically_plot_data, digits = 1, ordered = FALSE, legend.position = "right", text.size = 5) +
  labs(title = "Typically, studies that analyze preexisting observational datasets\n(such as the ALSPAC dataset) are...") +
  theme(
    title = element_text(size = 14),
    axis.text.y = element_text(size = 14),
    axis.text.x = element_text(size = 15),
    axis.title.x = element_blank(),
    legend.text = element_text(size = 13.5),
    legend.title = element_blank()
  )

# Create ecaw trustworthy and reproducible plot
# Calculate missing responses
ecaw_missing_text <-
  processed %>% 
  select(contains("ecaw_")) %>% 
  pivot_longer(
    cols = everything(),
    names_to = "item",
    values_to = "value"
    ) %>% 
  dplyr::filter(is.na(value) | value == "I don't understand the question") %>% 
  mutate(item = str_to_title(str_remove(item , "^[^_]*_"))) %>% 
  replace_na(list(value = "Missing")) %>% 
  group_by(item , value) %>% 
  summarise(
    n = n()
  ) %>% 
  ungroup() %>% 
  mutate(text = paste(value, "=", n))

# Prepare plot data
ecaw_plot_data <- 
  processed %>%
  select(contains("ecaw_")) %>% 
  # Transform not wanted response values to NA
  # likert::likert drops NA values silently
  # When var transformed to factor these values would be transformed to NA
  # Automatically but I try to be explicit
  mutate(
    across(
      everything(),
      ~ case_when(
        . == "I don't understand the question" ~ NA_character_,
        TRUE ~ .
      )
    )
  ) %>% 
  mutate(
    across(
      .fns = ~ factor(., levels =   c(
          "Much less",
          "Somewhat less",
          "About the same",
          "Somewhat more",
          "Much more"))
    )) %>% 
  rename(
    Reproducible = ecaw_reproducible,
    Trustworthy = ecaw_trustworthy
  )

@@

# Create figure
ecaw_plot <- 
  plot(ecaw_plot_data, digits = 1, ordered = FALSE, legend.position = "right", text.size = 5) +
  labs(title = "Compared to a typical study using preexisting observational data,\na study using an ECAW would be...") +
  theme(
    title = element_text(size = 14),
    axis.text.y = element_text(size = 14),
    axis.text.x = element_text(size = 15),
    axis.title.x = element_text(size = 15),
    legend.text = element_text(size = 13.5),
    legend.title = element_blank(),
    # legend.box.spacing = unit(0, "cm")
    legend.box.margin = margin(t = 0, r = 0, b = 0, l = -2, unit = "cm")
  )

@@

```{r include=FALSE}
confirmatory <- level_percentage(
  processed,
  method_confirmatory,
  c(
    "Never or almost never",
    "Sometimes",
    "About half the time",
    "Most of the time",
    "Always or almost always"),
  "I don't understand the question"
  ) %>% 
  filter(method_confirmatory != "Never or almost never") %>% 
  summarise(sum_percentage = sum(percentage)) %>% 
  pull(sum_percentage)

exploratory <- level_percentage(
  processed,
  method_exploratory,
  c(
    "Never or almost never",
    "Sometimes",
    "About half the time",
    "Most of the time",
    "Always or almost always"),
  "I don't understand the question"
  ) %>% 
  dplyr::filter(method_exploratory != "Never or almost never") %>% 
  summarise(sum_percentage = sum(percentage)) %>% 
  pull(sum_percentage)

preregistered <- level_percentage(
  processed,
  method_preregistered,
  c(
    "Never or almost never",
    "Sometimes",
    "About half the time",
    "Most of the time",
    "Always or almost always"),
  "I don't understand the question"
  )

blind <- level_percentage(
  processed,
  method_blind,
  c(
    "Never or almost never",
    "Sometimes",
    "About half the time",
    "Most of the time",
    "Always or almost always"),
  "I don't understand the question"
  )

script <- level_percentage(
  processed,
  method_script,
  c(
    "Never or almost never",
    "Sometimes",
    "About half the time",
    "Most of the time",
    "Always or almost always"),
  "I don't understand the question"
  )
```

@@

```{r methodPlot, warning=FALSE, fig.cap="Caption goes here...", fig.align="center", out.width="100%", fig.width=11, fig.height=4.5}
# Used methods figures
method_plot_data <-
  processed %>% 
    select(starts_with("method_")) %>% 
    pivot_longer(
      cols = everything(),
      names_to = "variable",
      values_to = "scale",
    ) %>%
    separate(variable, into = c("name_prefix", "type")) %>% 
    group_by(type) %>% 
    count(scale) %>% 
  ungroup() %>% 
    replace_na(list(scale = "Missing")) %>% 
    mutate(
      scale = factor(scale, levels = c(
          "Never or almost never",
          "Sometimes",
          "About half the time",
          "Most of the time",
          "Always or almost always",
          "Missing",
          "I don't understand the question")
            ),
      type = case_when(
        type == "blind" ~ "Blind the\ndata analyst",
        type == "confirmatory" ~ "Contain\nconfirmatory analysis",
        type == "exploratory" ~ "Contain\nexploratory analysis",
        type == "preregistered" ~ "Are preregistered",
        type == "script" ~ "Share analysis scripts",
      ),
      type = factor(type, levels = c(
        "Contain\nexploratory analysis",
        "Contain\nconfirmatory analysis",
        "Blind the\ndata analyst",
        "Share analysis scripts",
        "Are preregistered"
      ))
      ) %>% 
  group_by(type) %>%
  tidyr::complete(scale, fill = list(n = 0)) %>% 
  mutate(
    percentage = n / sum(n)
  ) %>% 
  ungroup()

@@

method_plot_data %>% 
  # filter(scale %ni% c("Missing", "I don't understand the question")) %>% 
  ggplot() +
  aes(
    x = percentage,
    y = type,
    fill = scale
  ) +
  geom_bar(position = "stack", stat = "identity") +
  scale_x_continuous(
    limits = c(0, 1),
    labels = scales::label_percent()
  ) +
  scale_fill_viridis(option = "A", discrete = TRUE) +
  labs(
    x = "Percentage",
    fill = "Response",
    title = "The studies using preexisting observational data that I am involved in..."
  ) +
  papaja::theme_apa() +
  theme(
    title = element_text(size = 17),
    axis.title.y = element_blank(),
    axis.text = element_text(size = 15),
    axis.title.x = element_text(size = 17),
    legend.text = element_text(size = 14),
    legend.title = element_blank(),
    legend.box.margin = margin(t = 0, r = 0, b = 0, l = -0.5, unit = "cm")
  ) 
```

```{r include=FALSE, warning=FALSE, message=FALSE}
willing <- support_percentage(
  processed,
  alspac_less_willing,
  c(
    "Strongly disagree",
    "Somewhat disagree",
    "Neither agree nor disagree",
    "Somewhat agree",
    "Strongly agree"),
  c(
    "I don't understand the question",
    "Unsure"
    )
  )

opt <- support_percentage(
  processed,
  alspac_opt_in,
  c(
    "Strongly disagree",
    "Somewhat disagree",
    "Neither agree nor disagree",
    "Somewhat agree",
    "Strongly agree"),
  c(
    "I don't understand the question",
    "Unsure"
    )
  )

study <- support_percentage(
  processed,
  alspac_study,
  c(
    "Strongly disagree",
    "Somewhat disagree",
    "Neither agree nor disagree",
    "Somewhat agree",
    "Strongly agree"),
  c(
    "I don't understand the question",
    "Unsure"
    )
  )

prefer <- support_percentage(
  processed,
  alspac_prefer_ecaw,
  c(
    "Strongly disagree",
    "Somewhat disagree",
    "Neither agree nor disagree",
    "Somewhat agree",
    "Strongly agree"),
  c(
    "I don't understand the question",
    "Unsure"
    )
  )
```

@@

```{r alspacPlot, warning=FALSE, fig.cap="Caption goes here...", fig.align="center", out.width="100%", fig.width=14, fig.height=4, message=FALSE}
# Check if there is any missing and don't understand responses separately
alspac_missing_text <-
  processed %>% 
  select(contains("alspac_")) %>% 
  pivot_longer(
    cols = everything(),
    names_to = "item",
    values_to = "value"
    ) %>% 
  dplyr::filter(is.na(value) | value == "I don't understand the question" | value == "Unsure") %>% 
  mutate(item = str_to_title(str_remove(item , "^[^_]*_"))) %>% 
  replace_na(list(value = "Missing")) %>% 
  group_by(item , value) %>% 
  summarise(
    n = n()
  ) %>% 
  ungroup() %>% 
  mutate(text = paste(value, "=", n))

# Prepare plot data
alspac_plot_data <- 
  processed %>%
  select(contains("alspac_")) %>% 
  # Transform not wanted response values to NA
  # likert::likert drops NA values silently
  # When var transformed to factor these values would be transformed to NA
  # Automatically but I try to be explicit
  mutate(
    across(
      everything(),
      ~ case_when(
        . %in% c("I don't understand the question", "Unsure") ~ NA_character_,
        TRUE ~ .
      )
    )
  ) %>% 
  mutate(
    across(
      .fns = ~ factor(., levels = c(
          "Strongly disagree",
          "Somewhat disagree",
          "Neither agree nor disagree",
          "Somewhat agree",
          "Strongly agree")
          )
      )
    ) %>% 
  rename(
    `If ALSPAC required that I use an ECAW, I would be less willing to use their data in my research` = alspac_less_willing,
    `If ALSPAC ran a study on ECAWs, I would opt-in.` = alspac_opt_in,
    `ALSPAC should run a study on ECAWs.` = alspac_study,
    `I would prefer using an ECAW than using typical preregistration` = alspac_prefer_ecaw
  )

@@

# Create figure
plot(alspac_plot_likert_data, digits = 1, ordered = TRUE, group.order = names(alspac_plot_data), text.size = 6) +
  labs(title = "Thinking about a study you may run with ALSPAC data (or one that you have recently run)...") +
  theme(
    title = element_text(size = 14),
    axis.text.y = element_text(size = 13),
    axis.text.x = element_text(size = 15),
    axis.title.x = element_text(size = 15),
    legend.text = element_text(size = 13),
    legend.title = element_blank()
  )
```

>>>>

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
library("papaja")
library(tidyverse)
library(here)
library(kableExtra)
library(gt)
library(patchwork)
# The developer version is needed
library(likert)
library(knitr)
library(viridis)
# r_refs("references.bib")
```

```{r customScripts}
# Source R scripts
r_scripts <- list.files(here::here("R/"), full.names = TRUE)
purrr::walk(r_scripts, source)
```

```{r preprocessing, include=FALSE}
# loads source data, performs preprocessing, saves raw datafiles
xfun::Rscript_call(
  rmarkdown::render,
  list(
    input = here::here("preprocessing", "ecaw_source_raw_preprocessing.Rmd"),
    output_format = "html_document"
  )
)

# loads raw data, performs preprocessing, saves processed datafiles
xfun::Rscript_call(
  rmarkdown::render,
  list(
    input = here::here("preprocessing", "ecaw_raw_processed_preprocessing.Rmd"),
    output_format = "html_document"
  )
)
```

```{r loadingData}
processed <- read_csv(here::here("data/processed/ecaw_processed_data.csv"))
```

```{r analysisPreferences}
# Seed for random number generation
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

@@

```{r languageDescriptives, include=FALSE}
# List of used programming languages and their counts text
language_text <- 
  processed %>%
  rename(lang = programming_language) %>%
  separate_rows(lang, sep = ",") %>%
  count(lang) %>%
  filter(!is.na(lang)) %>%
  arrange(desc(n)) %>%
  # Removing irrelevant
  filter(lang != "irrelevant") %>%
  mutate(lang_n = glue::glue("{lang} (n = {n})")) %>%
  pull(lang_n) %>%
  glue::glue_collapse(., sep = ", ", last = ", and ")
```

```{r concernedDescriptives, include=FALSE}
# Calculate concerned percentage
concerned <- support_percentage(
  processed,
  concerned,
  c(
    "very much less concerned",
    "less concerned",
    "somewhat less concerned",
    "as concerned as a typical researcher in my field",
    "somewhat more concerned",
    "more concerned",
    # This level label starts with a capital letter
    "Very much more concerned"
  )
)
```

@@

```{r typicallyEcawDescriptives, include=FALSE}
# Calculate typical trustworthiness and reproducibility ratings
typically_trustworthy <- support_percentage(
  processed,
  typically_trustworthy,
  c(
    "Strongly disagree",
    "Somewhat disagree",
    "Neither agree nor disagree",
    "Somewhat agree",
    "Strongly agree"
  ),
  "I don't understand the question"
)

typically_reproducible <- support_percentage(
  processed,
  typically_reproducible,
  c(
    "Strongly disagree",
    "Somewhat disagree",
    "Neither agree nor disagree",
    "Somewhat agree",
    "Strongly agree"
  ),
  "I don't understand the question"
)

# Calculate ecaw trustworthiness and reproducibility ratings
ecaw_trustworthy <- support_percentage(
  processed,
  ecaw_trustworthy,
  c(
    "Much less",
    "Somewhat less",
    "About the same",
    "Somewhat more",
    "Much more"
  ),
  "I don't understand the question"
)

ecaw_reproducible <- support_percentage(
  processed,
  ecaw_reproducible,
  c(
    "Much less",
    "Somewhat less",
    "About the same",
    "Somewhat more",
    "Much more"
  ),
  "I don't understand the question"
)
```

```{r typicallyEcawPlot, warning=FALSE, fig.cap="Caption goes here...", fig.align="center", out.width="100%", fig.width=12.8, fig.height=6, fig.path='figs/', dev=c('png', 'pdf')}
# Create typically trustworthy and reproducible plot
# Check if there is any missing and don't understand responses separately
# processed %>%
#   select(contains("typically_")) %>%
#   dplyr::filter(is.na(typically_reproducible) | is.na(typically_trustworthy) | typically_reproducible == "I don't understand the question" | typically_trustworthy == "I don't understand the question") %>%
#   nrow()

# Prepare plot data
typically_plot_data <-
  processed %>%
  select(contains("typically_")) %>%
  mutate(
    across(
      .fns = ~ factor(., levels = c(
        "Strongly disagree",
        "Somewhat disagree",
        "Neither agree nor disagree",
        "Somewhat agree",
        "Strongly agree"
      ))
    )
  ) %>%
  rename(
    Reproducible = typically_reproducible,
    Trustworthy = typically_trustworthy
  )

@@

# Create figure
typically_plot <-
  plot(typically_plot_data, digits = 1, ordered = FALSE, legend.position = "right", text.size = 5) +
  scale_y_continuous(labels = c("100%", "50%", "0%", "50%", "100%"), limits = c(-105, 105)) +
  labs(title = "Typically, studies that analyze preexisting observational datasets\n(such as the ALSPAC dataset) are...") +
  theme(
    title = element_text(size = 14),
    axis.text.y = element_text(size = 14),
    axis.text.x = element_text(size = 15),
    axis.title.x = element_blank(),
    legend.text = element_text(size = 13.5),
    legend.title = element_blank()
  )

# Create ecaw trustworthy and reproducible plot
# Calculate missing responses
ecaw_missing_text <-
  processed %>%
  select(contains("ecaw_")) %>%
  pivot_longer(
    cols = everything(),
    names_to = "item",
    values_to = "value"
  ) %>%
  dplyr::filter(is.na(value) | value == "I don't understand the question") %>%
  mutate(item = str_to_title(str_remove(item, "^[^_]*_"))) %>%
  replace_na(list(value = "Missing")) %>%
  group_by(item, value) %>%
  summarise(
    n = n()
  ) %>%
  ungroup() %>%
  mutate(text = paste(value, "=", n))

# Prepare plot data
ecaw_plot_data <-
  processed %>%
  select(contains("ecaw_")) %>%
  # Transform not wanted response values to NA
  # likert::likert drops NA values silently
  # When var transformed to factor these values would be transformed to NA
  # Automatically but I try to be explicit
  mutate(
    across(
      everything(),
      ~ case_when(
        . == "I don't understand the question" ~ NA_character_,
        TRUE ~ .
      )
    )
  ) %>%
  mutate(
    across(
      .fns = ~ factor(., levels = c(
        "Much less",
        "Somewhat less",
        "About the same",
        "Somewhat more",
        "Much more"
      ))
    )
  ) %>%
  rename(
    Reproducible = ecaw_reproducible,
    Trustworthy = ecaw_trustworthy
  )

@@

# Create figure
ecaw_plot <-
  plot(ecaw_plot_data, digits = 1, ordered = FALSE, legend.position = "right", text.size = 5) +
  scale_y_continuous(labels = c("100%", "50%", "0%", "50%", "100%"), limits = c(-105, 105)) +
  labs(title = "Compared to a typical study using preexisting observational data,\na study using an ECAW would be...") +
  theme(
    title = element_text(size = 14),
    axis.text.y = element_text(size = 14),
    axis.text.x = element_text(size = 15),
    axis.title.x = element_blank(),
    legend.text = element_text(size = 13.5),
    legend.title = element_blank(),
    legend.box.margin = margin(t = 0, r = 0, b = 0, l = -2.3, unit = "cm")
  )

@@

```{r methodDescriptives, include=FALSE}
confirmatory <- level_percentage(
  processed,
  method_confirmatory,
  c(
    "Never or almost never",
    "Sometimes",
    "About half the time",
    "Most of the time",
    "Always or almost always"
  ),
  "I don't understand the question"
) %>%
  filter(method_confirmatory != "Never or almost never") %>%
  summarise(sum_percentage = sum(percentage)) %>%
  pull(sum_percentage)

exploratory <- level_percentage(
  processed,
  method_exploratory,
  c(
    "Never or almost never",
    "Sometimes",
    "About half the time",
    "Most of the time",
    "Always or almost always"
  ),
  "I don't understand the question"
) %>%
  dplyr::filter(method_exploratory != "Never or almost never") %>%
  summarise(sum_percentage = sum(percentage)) %>%
  pull(sum_percentage)

preregistered <- level_percentage(
  processed,
  method_preregistered,
  c(
    "Never or almost never",
    "Sometimes",
    "About half the time",
    "Most of the time",
    "Always or almost always"
  ),
  "I don't understand the question"
)

blind <- level_percentage(
  processed,
  method_blind,
  c(
    "Never or almost never",
    "Sometimes",
    "About half the time",
    "Most of the time",
    "Always or almost always"
  ),
  "I don't understand the question"
)

script <- level_percentage(
  processed,
  method_script,
  c(
    "Never or almost never",
    "Sometimes",
    "About half the time",
    "Most of the time",
    "Always or almost always"
  ),
  "I don't understand the question"
)
```

@@

```{r methodPlot, warning=FALSE, fig.cap="Caption goes here...", fig.align="center", out.width="100%", fig.width=11, fig.height=4.5, fig.path='figs/', dev=c('png', 'pdf')}
# Used methods figures
method_plot_data <-
  processed %>%
  select(starts_with("method_")) %>%
  pivot_longer(
    cols = everything(),
    names_to = "variable",
    values_to = "scale",
  ) %>%
  separate(variable, into = c("name_prefix", "type")) %>%
  group_by(type) %>%
  count(scale) %>%
  ungroup() %>%
  replace_na(list(scale = "Missing")) %>%
  mutate(
    scale = factor(scale, levels = c(
      "Never or almost never",
      "Sometimes",
      "About half the time",
      "Most of the time",
      "Always or almost always",
      "Missing",
      "I don't understand the question"
    )),
    type = case_when(
      type == "blind" ~ "Blind the data analyst",
      type == "confirmatory" ~ "Contain confirmatory analysis",
      type == "exploratory" ~ "Contain exploratory analysis",
      type == "preregistered" ~ "Are preregistered",
      type == "script" ~ "Share analysis scripts",
    ),
    type = factor(type, levels = c(
      "Contain exploratory analysis",
      "Contain confirmatory analysis",
      "Blind the data analyst",
      "Share analysis scripts",
      "Are preregistered"
    ))
  ) %>%
  group_by(type) %>%
  tidyr::complete(scale, fill = list(n = 0)) %>%
  mutate(
    percentage = n / sum(n)
  ) %>%
  ungroup()

@@

method_plot_data %>%
  # filter(scale %ni% c("Missing", "I don't understand the question")) %>%
  ggplot() +
  aes(
    x = percentage,
    y = type,
    fill = scale
  ) +
  geom_bar(position = "stack", stat = "identity") +
  scale_x_continuous(
    limits = c(0, 1),
    labels = scales::label_percent()
  ) +
  scale_fill_viridis(option = "A", discrete = TRUE) +
  scale_y_discrete(labels = function(x) str_wrap(x, width = 20)) +
  labs(
    title = "The studies using preexisting observational data that I am involved in..."
  ) +
  papaja::theme_apa() +
  theme(
    title = element_text(size = 17),
    axis.text = element_text(size = 15),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    legend.text = element_text(size = 14),
    legend.title = element_blank(),
    legend.box.margin = margin(t = 0, r = 0, b = 0, l = -0.5, unit = "cm")
  )
```

```{r aslpacDescriptives, include=FALSE, warning=FALSE, message=FALSE}
willing <- support_percentage(
  processed,
  alspac_less_willing,
  c(
    "Strongly disagree",
    "Somewhat disagree",
    "Neither agree nor disagree",
    "Somewhat agree",
    "Strongly agree"
  ),
  c(
    "I don't understand the question",
    "Unsure"
  )
)

opt <- support_percentage(
  processed,
  alspac_opt_in,
  c(
    "Strongly disagree",
    "Somewhat disagree",
    "Neither agree nor disagree",
    "Somewhat agree",
    "Strongly agree"
  ),
  c(
    "I don't understand the question",
    "Unsure"
  )
)

study <- support_percentage(
  processed,
  alspac_study,
  c(
    "Strongly disagree",
    "Somewhat disagree",
    "Neither agree nor disagree",
    "Somewhat agree",
    "Strongly agree"
  ),
  c(
    "I don't understand the question",
    "Unsure"
  )
)

prefer <- support_percentage(
  processed,
  alspac_prefer_ecaw,
  c(
    "Strongly disagree",
    "Somewhat disagree",
    "Neither agree nor disagree",
    "Somewhat agree",
    "Strongly agree"
  ),
  c(
    "I don't understand the question",
    "Unsure"
  )
)
```

@@

```{r alspacPlot, warning=FALSE, message=FALSE, fig.cap="Caption goes here...", fig.align="center", out.width="100%", fig.width=14, fig.height=5, fig.path='figs/', dev=c('png', 'pdf')}
# Check if there is any missing and don't understand responses separately
alspac_missing_text <-
  processed %>%
  select(contains("alspac_")) %>%
  pivot_longer(
    cols = everything(),
    names_to = "item",
    values_to = "value"
  ) %>%
  dplyr::filter(is.na(value) | value == "I don't understand the question" | value == "Unsure") %>%
  mutate(item = str_to_title(str_remove(item, "^[^_]*_"))) %>%
  replace_na(list(value = "Missing")) %>%
  group_by(item, value) %>%
  summarise(
    n = n()
  ) %>%
  ungroup() %>%
  mutate(text = paste(value, "=", n))

# Prepare plot data
alspac_plot_data <-
  processed %>%
  select(contains("alspac_")) %>%
  # Transform not wanted response values to NA
  # likert::likert drops NA values silently
  # When var transformed to factor these values would be transformed to NA
  # Automatically but I try to be explicit
  mutate(
    across(
      everything(),
      ~ case_when(
        . %in% c("I don't understand the question", "Unsure") ~ NA_character_,
        TRUE ~ .
      )
    )
  ) %>%
  mutate(
    across(
      .fns = ~ factor(., levels = c(
        "Strongly disagree",
        "Somewhat disagree",
        "Neither agree nor disagree",
        "Somewhat agree",
        "Strongly agree"
      ))
    )
  ) %>%
  rename(
    `If ALSPAC required that I use an ECAW, I would be less willing to use their data in my research` = alspac_less_willing,
    `If ALSPAC ran a study on ECAWs, I would opt-in.` = alspac_opt_in,
    `ALSPAC should run a study on ECAWs.` = alspac_study,
    `I would prefer using an ECAW than using typical preregistration` = alspac_prefer_ecaw
  )

@@

# Create figure
plot(alspac_plot_likert_data, digits = 1, ordered = TRUE, text.size = 6) +
  # TODO: have to check if this is not messing with the results
  scale_x_discrete(labels = function(x) str_wrap(x, width = 30), limits = rev(names(alspac_plot_data))) +
  scale_y_continuous(labels = c("100%", "50%", "0%", "50%", "100%"), limits = c(-105, 105)) +
  labs(title = "Thinking about a study you may run with ALSPAC data (or one that you have recently run)...") +
  theme(
    axis.title.x = element_blank(),
    title = element_text(size = 15.5),
    axis.text.y = element_text(size = 14),
    axis.text.x = element_text(size = 16),
    legend.text = element_text(size = 15),
    legend.title = element_blank()
  )
```
